\chapter{A guessing game}
We stop your usual broadcast to present a guessing game for you.\\
The reason why most Olympiad givers find PnC hard is as they cannot decide which strategy to use. While the other strategies we are yet to study have a signature framing which you can notice with practice, the most difficult to discern and the most fundamental methods of counting by far are: constructive counting(ch-2), complementary counting(ch-3) and casework(ch-3).  In the next few questions you have to first guess what to use. Then solve the question and check if your guess was correct. \\
I have not provided points or hints for these questions as they are more like a test then an exercise.\\
Also note that the questions were randomized by a computer, so don't look for patterns. You'll not find them.
\begin{xcb}{Exercises}
\begin{enumerate}
\item (IOQM 2021) Find the number of maps $f:\{1, 2, 3\} \rightarrow \{1, 2, 3, 4, 5\}$ such that $f (i) \leq f (j)$ whenever $i < j$ .
\item Consider all 3 element subsets of ${1,2,3 \dots 298, 299,300}$ What number of the subsets have the sum of elements divisible by 3?
\item How many three term increasing GPs can be obtained from the GP $1, 2, 2^2, \dots 2^n$
\item In how many ways can five cards labeled $A,B,C,D,E$ be rearranged so that no card is more than one away from its original position?(Not moving the cards is also a valid arrangement)
\item (ARML 2011) The six sides of a convex hexagon $A_1A_2A_3A_4A_5A_6$ are colored red. Each of the diagonals of the hexagon is colored either red or blue. If $N$ is the number of coloring's such that every triangle $A_iA_jA_k$ , where $1 \leq i < j < k \leq 6$ , has at least one red side, find the sum of the digits of N .
\item (AIME 2010) Let $N$ be the number of ways to write $2010$ in the form $2010 = a_3 \cdot 10^3 + a_2 \cdot 10^2 + a_1 \cdot 10 + a_0$, where the $a_i$'s are integers, and $0 \le a_i \le 99$. An example of such a representation is $1\cdot 10^3 + 3\cdot 10^2 + 67\cdot 10^1 + 40\cdot 10^0$. Find $N$.
\item (Purple Comet 2013) How many four-digit positive integers have exactly one digit equal to $1$ and exactly one digit equal to $3$?
\item (Pascal 2005) A number is called special if each of its digits is less than the digit to the left. For example, 5420 is special. How many special numbers are there between 200 and 700?
\item How many three element subsets of ${1,2,3 \dots 18, 19, 20}$ have there product divisible by 4?
\item A cube is constructed from 4 white unit cubes and 4 blue unit cubes. How many different ways are there to construct the $2*2*2$ cube using these smaller cubes? (Two constructions are considered the same if one can be rotated to match the other.)
\item (IOQM 2020) Three couples sit for a photograph in $2$ rows of three people each such that no couple is sitting in the same row next to each other or in the same column one behind the other. How many arrangements are possible?
\item Let A denote a subset of ${1,11,21, \dots 531,541,551}$, where any two elements of A do not add up to 552. What is the maximum number of elements in A?
\item (AIME 2005) A game uses a deck of n dierent cards, where n is an integer and $n \geq 6$. The number of possible sets of 6 cards that can be drawn from the deck is 6 times the number of possible sets of 3 cards that can be drawn. Find n.
\item (Fermat 2007) What is the number of 3-digit positive integers a such that both a and 2a have only even digits?
\item Find how many committees(of any number of people more than 1) with a chairman can be chosen from a set of n persons.
\item (AIME 2014) Let the set $S = \{P_1, P_2, \dots, P_{12}\}$ consist of the twelve vertices of a regular $12$-gon. A subset $Q$ of $S$ is called "communal" if there is a circle such that all points of $Q$ are inside the circle, and all points of $S$ not in $Q$ are outside of the circle. How many communal subsets are there? (Note that the empty set is a communal subset.)
\item (Cayley 2009) How many integers n are there with the property that the product of the digits of $n$ is $0$, where $5000 \leq n \leq 6000$?
\item How many 4 digits exist such that the thousands digit is 1 and the number has exactly two equal digits?
\item (Fermat 2010) A gumball machine that randomly dispenses one gumball at a time contains 13 red, 5 blue, 1 white, and 9 green gumballs. What is the least number of gumballs that Wally must buy to guarantee that he receives 3 gumballs of the same colour?
\item (2004 AIME) An integer is called snakelike if its decimal representation $a_1a_2a_3 \dots a_k$ satisfies $a_i < a_{i+1}$  if i is odd and $a_i > a_{i+1}$ if i is even. How many snakelike integers between 1000 and 9999 have four distinct digits?
\item (AIME 2014) An urn contains $4$ green balls and $6$ blue balls. A second urn contains $16$ green balls and $N$ blue balls. A single ball is drawn at random from each urn. The probability that both balls are of the same color is $0.58$. Find $N$
\item (1992 AIME) A positive integer is called ascending if, in its decimal representation, there are at least two digits and each digit is less than any digit to its right. How many ascending positive integers are there?
\item (AIME 2018) Find the number of permutations of $1, 2, 3, 4, 5, 6$ such that for each k with $1 \leq k \leq 5$ at least one of the first k terms of the permutation is greater than k.
\item (IOQM 2021) In how many ways can four married couples sit in a merry-go-round with identical seats such that men and women occupy alternate seats and no husband seats next to his wife?
\item Call a positive integer an uphill integer if every digit is strictly greater than the previous digit. For example, 1357, 89, and 5 are all uphill integers, but 32, 1240, and 466 are not. How many uphill integers are divisible by 15?
\item For a particular peculiar pair of dice, the probabilities of rolling $1, 2, 3, 4, 5,6$, on each die are in the ratio $1 : 2 : 3 : 4 : 5 : 6$ What is the probability of rolling a total of 7 on the two dice?
\item (IOQM 2023) Unconventional dice are to be designed such that the six faces are marked with numbers from 1 to 6 with 1 and 2 appearing on opposite faces. Further, each face is colored either red or yellow with opposite faces always of the same color. Two dice are considered to have the same design if one of them can be rotated to obtain a dice that has the same numbers and colors on the corresponding faces as the other one. Find the number of distinct dice that can be designed.
\item A restaurant has six appetizers, five main courses, and four deserts to choose from its menu. How many possible dinners are there if a main course is required but appetizers and deserts are not?
\item (AIME 2010) Dave arrives at an airport which has twelve gates arranged in a straight line with exactly $100$ feet between adjacent gates. His departure gate is assigned at random. After waiting at that gate, Dave is told the departure gate has been changed to a different gate, again at random. Let the probability that Dave walks $400$ feet or less to the new gate be a fraction $\frac{m}{n}$, where $m$ and $n$ are relatively prime positive integers. Find $m+n$.
\item (Cayley 2010) What is the number of 3-digit positive integers that have exactly one even digit?
\end{enumerate}
\end{xcb}
